% This will be my acknowledgements section
% it will be very witty and incisive!

%\newpage
%\chapter*{Acknowledgements}

\newpage
%\makeatletter
%\vspace*{-50\p@} % makeschapterhead starts with a vspace of 50pt
\chapter*{Acknowledgements}
%\@makeschapterhead{Acknowledgements}
%\makeatother
I was attracted to the field of \gw research due to the promise that we would be entering exciting
times with several large scale projects bringing in unprecedented amounts of new data. Given this
the discovery of \gws would be just round the corner, opening up \gw astronomy for real. Little did
I know that this has been exactly what's been said for around 30 years! Despite this I do actually
believe that I've entered the field at a prime time and finding \gws is just around, if not the
first corner, then the next one. I've even got money on it! I'd therefore like to thank the
people who have encouraged me into pursuing this, sometimes frustrating, but hopefully ultimately
very rewarding, career path.

Three years ago I was made very welcome at the University of Glasgow after moving up from London.
Much of that was down to the very friendly people in the two research groups I became a member of:
the Institute for Gravitational Research (IGR) and the Astronomy \& Astrophysics (A\&A) group. When
I started my PhD the \gw data analysis group at Glasgow was rather small, with one faculty member
and three students (including me) nestled in one small office. Being such a small group I can
imagine it was a risk taking me on as a student, so I'm therefore very grateful to my supervisor
Graham Woan and IGR leader Jim Hough for taking that risk and all their help since. I'm also
indebted to my original officemates R\'ejean and Martin. Without R\'ejean's work to build on my PhD
would have taken a very different path, and Martin was always helpful with Matlab and signal
processing queries. I should also thank PPARC for funding me through my PhD.

Since then us data analysts have been marching forward attracting more and more people. Although
both R\'ejean and Martin have moved on (leaving me with the daunting task of filling their shoes and
being the most {\it experienced} student!) the new guys to come in have been equally good. I'd
therefore like to say a big thanks to John and Jen for being brilliant officemates and making ours
the best office in the Kelvin building (keep up the good work with James). 

With the official business done I'll put you out of your misery in finding out who else has made it
into the acknowledgement section and who hasn't. Going back to when I started my PhD I was quickly
drawn into the close knit group of A\&A students. So thanks to: Kirsty - for many nights of vodka
and diet Coke; Ross - for his moments of extreme outrage; Iain - for providing an unmitigated
barrage of sarcasm and web links; Brendan - for full on gingerness; Brian, Gail, Eduard, Hugh etc -
for football; Graeme and Norman - for all things computer based; and everyone else for being so
nice, aaahh.

In the last year there have been many people who've taken on the task of stopping me finishing my
PhD, so here I give them their dues. Bob has done a fine job deflecting anyone else's misdemeanours
by providing a reputation no-one can match. Helen has valiantly dragged us to the pub at any and
every sunny opportunity, with Richard and Neil only to happy to come along. A big seven out of ten
goes to Rav, who by returning to Glasgow got me back in the pub watching football where I belong.
Mandy also deserves thanks for providing a much needed distraction from the horrors of work. 

My friends and family south of the border are also deserving a large amount of thanks. Thank you Mum
and Dad, Rebecca - for giving me someone to look up to, Gemma and Andrew. Colin, Neil, Jim, Steve,
Justin and all the others with whom I got through Sandringham School have been great friends over
the years. James, Mike, Rich, Eve and all my other friends from UCL gave me a great time as an
undergraduate and continue to be good friends. All of whom have provided great encouragement for,
and more often a refuge from, my work.

Thanks should also be extended to a great many members of the LIGO Scientific Collaboration (LSC),
for making collaboration meetings a sometimes enjoyable experience, and for all the help and
advice that's been extended to me.

As ever there are people who've had to be left out here, but a great big thanks to you all wherever
you are.